Bayesian Evidence Synthesis:opioid crisis

### 2020-02-02 article_4/newpaper.tex
latex major revision done in the middle of Sunday

The number of overdoses is our ultimate interest of estimation. To achive the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation. The posterior belief is coming from both parts of the information. Following equation express the idea in a mathmatical form:  
\begin{equation}
\label{bayes_thm}
\left.\begin{aligned}
p(\theta| y ) = \frac{ p(\theta, y)   }{ p(y)} = \frac{ p(\theta) p(y| \theta)}{ p(y)} \propto p(\theta) p(y| \theta) \end{aligned}\right.
\end{equation}
where $\theta$ represents the target of estimation, the total overdoses and $y$ represents the collected data set providing relevant samples of the target variable. In Bayesian concepts, the variable of interest $\theta$ is a fixed number but the number is yielded from a certain distribution; That is, $\theta$ is considered to be a sample of a random variable.

The problem, however, is that the data set $y$ is not available. Hence the technique is to approach the estimation in a indirect fashion. Figure \cite{simple_drawing} shows how we estimate the target variable. 
\\

@@

\begin{equation}
\label{ambulance}
\left.\begin{aligned}
x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
\text{		ambulance call-outs model}
\end{equation}

The total overdoses need to be modeled. The simplest conceptual model is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$. \cite{Irvine:modelling} Denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$. It is assumed that the total overdose $O_t$ follows Poisson distribution where the population of the region of interest is $N$. 
[Give a graphical model representation? state that(?)]

@@

\normalsize 
The first simulation simplifies the assumptions of variables as much as possible; We assumed $N= 10000, n_{A}=1000$. Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity. The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; The total population size for a region, N, could vary over time or it can be staritified for a better realization of the real world. We vary the survey size  as $n_{A}=100$ or $n_{A}=10000$.  \\

[brief explaination of MCMC and related topics]
\subsubsection{Likelihood}

@@

\newpage
\subsubsection{Posterior Predictive Check}

Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$. It is shown that the posterior predictive estimates of $U_t$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $O_t$. It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11). Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; For months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\

@@

\normalsize 
Figure \ref{contam_ot} is the boxplots of posterior samples where $p_A$ is contaminated from the survey data. It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots. This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); $O_t$ is generated by multiplying $U_t$ and the inverse of $p_A$ where $p_A$ is underestimated; Hence $\frac{1}{p_A}$ is overestimated which leads overestimation of $O_t$.
\begin{equation}
\label{ot.how.made}
\begin{aligned}
O_t = U_t \frac{1}{p_A}
\end{aligned}
\end{equation}
Figure \ref{over_ot} shows the opposite case. Overestimation of $p_A$ leads underestimation of the inverse of $p_A$ which causes underestimation of $O_t$. From both figures it is seen that the bias increases as the estimated values and the actual values get large.\\

@@

@@

Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that the none of the contaminations of $p_A$ leads an effect  $U_t$. [Explain why is this]\\

\begin{figure}[htb]
	\centering
	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-x_a.png}\label{under_xa}}
	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
	\caption[two early result box plots:xa]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $x_A$ values are shown as red dots.}
	\label{contam_xt}
\end{figure}
Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots. However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900. [I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); Only the ambulance model has an effect on the overdose model.]

>>>>

The number of overdoses is our ultimate interest of estimation. To achive the goal in context of Bayesian statistics, we use our prior belief and available data set about the target of estimation. The posterior belief is coming from both parts of the information. Following equation express the idea in a mathmatical form:  
\begin{equation}
\label{bayes_thm}
\left.\begin{aligned}
p(\theta| y ) = \frac{ p(\theta, y)   }{ p(y)} = \frac{ p(\theta) p(y| \theta)}{ p(y)} \propto p(\theta) p(y| \theta) \end{aligned}\right.
\end{equation}
where $\theta$ represents the target of estimation, the total overdoses and $y$ represents the collected data set providing relevant samples of the target variable. In Bayesian concepts, the variable of interest $\theta$ is a fixed number but the number is yielded from a certain distribution; That is, $\theta$ is considered to be a sample of a random variable. The problem, however, is that the data $y$ is not available. We can have our prior belief about the total overdoses but it is not possible to obtain any data. Hence it is needed to approach the estimation in a indirect fashion. Figure \ref{simple_draw} shows an example of how the estimation can be proceeded. [attempt: Give a graphical model representation? state that(?)]
\begin{figure}[htb]
	\centering
	\includegraphics[width=0.5\linewidth]{Figures/simple_drawing}
	\caption{Example of estimating the total overdoses indirectly given two data sets: ambulance attended overdoses and survey data. N is the population size}
	\label{simple_draw}
\end{figure}
\\

@@

\begin{equation}
\label{ambulance}
\left.\begin{aligned}
x_{A} \sim Bin(n_{A},p_{A})\end{aligned}\right.
\text{ambulance call-outs model}
\end{equation}

The total overdoses need to be modeled. The simplest conceptual model is to take an underlying log-rate $z_t$ that is independent and identically distributed across months according to a normal distribution with mean $\mu$ and variance $\sigma^2$. \cite{Irvine:modelling} Denote $\lambda_{t}^{OD}$ the rate of overdose at time $t$. It is assumed that the total overdose $O_t$ follows Poisson distribution where the population of the region of interest is $N$.

@@

\normalsize 
The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$. Note that the population size, N,  is unknown in reality but it is assumed known for the investigator here for simplicity as in figure \ref{simple_draw}. The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality. The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.  \\

\subsubsection{Markov chain Monte Carlo}
The achievement of the simulation was done by Markov chain Monte Carlo (MCMC). The technique is a popular method to obtain posterior samples from distributions where analytic forms do not exist or are hard to be derived. [Further explanation is going to be added.] [attempt:brief explaination of MCMC and related topics]\\

\subsubsection{Likelihood}

@@

\newpage
\subsubsection{Posterior Predictive Check}
Posterior predictive checking is a model validation technique that we simulate some replicated data under the fitted model then compare the new data to the observed data. If the model fits, then replicated data generated under the model should look similar to observed data. To put it another way, the observed data should look plausible under the posterior predictive distribution. This is really a self-consistency check: an observed discrepancy can be due to model misfit or chance. \cite{bda_galman}
[brief explannation of what is PPC]\\

Figure \ref{ppc_ut} is the boxplot of posterior predictive samples of $U_t$. It is shown that the posterior predictive estimates of $U_t$ is failry accurate with the same two reasons regarding the accuracy of the posterior distribution of $O_t$. It is more obvious here that the range of the boxplots from higher $O_t$ values (t=1, 7, 11) is wider than the other ranges of the boxplots from smaller estimates of $O_t$ (all t values but 1,7, 11). Notice that the relative ranges of figure \ref{ppc_ut} follow the ones of figure \ref{pst_ot}; for months where $U_t$ values are higher, the values of $O_t$ are also higher than the average (t=1, 7, 11)\\

@@

\normalsize 
Figure \ref{contam_ot} is the boxplots of posterior samples where $p_A$ is contaminated from the survey data. It is seen that the underestimation of $p_A$ (\ref{under_ot}) leads to an overestimation of $O_t$ as the boxplots are above the red dots. This is justifiable considering the given data sets (likelihoods) and the relationship between the two models (\ref{over_amb}); given the likelihood of the ambulance attended overdoses, $U_t$, a value of $O_t$ is an unknown parameter of the binomial distribution, $Bin(O_t, p_A)$. Let $o_t$, $u_t$ are samples from $O_t$ and the likelihood value of $U_t$ respectively. Then $o_t$ is proportional to the quantity generated by multiplying $u_t$ and the inverse of $\hat{p}_{A}$ where $\hat{p}_{A}$ is underestimated estimator of $p_A$; hence $\frac{1}{\hat{p}_{A}}$ is overestimated which leads overestimation of $O_t$.
\begin{equation}
\label{ot.how.made}
\begin{aligned}
o_t \propto u_t \frac{1}{\hat{p}_{A}}
\end{aligned}
\end{equation}
Figure \ref{over_ot} shows the opposite case. Overestimation of $p_A$ leads underestimation of the inverse of $p_A$ which causes underestimation of $O_t$. From both figures it is seen that the bias increases as the estimated values and the actual values get large.\\

@@

Figure \ref{contam_ut} are the boxplots of posterior predictive samples where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that none of the contaminations of $p_A$ leads an contamination effect on $U_t$ but only affects the $O_t$ estimation.\\

The possible explanation is that the contaminations of  $p_A$ and $O_t$ (from $p_A$) cancel out the bias component as combined so that $U_t$ has no bias. Notice that $U_t$ is a likelihood (data set) of $O_t$ estimation. Hence it plays the role of $y$ in Bayes theorem (\ref{bayes_thm}) and the likelihood, $U_t$, is used to obtain the biased posterior samples of $O_t$ with the contaminated estimates of $p_A$. Here the pair of $(O_t, P_A)$ are both biased and the pair is fitted by the likelihood, $U_t$. To obtain the posterior predictive samples, MCMC algorithm help obtain posterior samples of biased $O_t$ and then the samples will be combined with biased $p_A$ such that the new samples of $\tilde{U_t}$ from $p(\tilde{U_t}|{U_t})$ is obtained.  The result also matches with our intution. Posterior predictive checkes use the existing data points twice; firstly it uses the data to obtain the posterior samples of paramaters, and secondly the samples are used to produce posterior predictive samples. Given the fact that there could be an overfitting issue due to using the data twice, the result shown here seems reasonable. 
\\

@@

Figure \ref{contam_xt}  are the boxplots of posterior predictive samples of $x_A$ where $p_A$ is underestimated and overestimated respectively from the survey data. It is seen that none of the contaminations of $p_A$ leads an effect of contamination on $x_A$; the actual simulated points (red dots) are close to the medians from the two boxplots. However, notice that the range of the estimated values are different between the two boxplots; the median from Figure \ref{under_xa} is around 600 whereas the median from Figure \ref{over_xa} is around 900. [I think this should go to the previous of the previous paragraph: This is because overdose model (\ref{overdose}) does not affect the resulf from the ambulance model (\ref{ambulance}); only the ambulance model has an effect on the overdose model.]

\begin{figure}[h]
	\centering
	\subfloat[ppc of $x_A$: underestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_underestimated-x_a.png}\label{under_xa}}
	\subfloat[ppc of $x_A$: overestimated $p_A$]{\includegraphics[width=.50\columnwidth]{early_contamination_overestimated-x_a.png}\label{over_xa}}
	\caption[two early result box plots:xa]{Boxplot of posterior samples of $x_A$ (1000 samples) where survey data is contaminated.  The actual data point of simulated $x_A$ values are shown as red dots.}
	\label{contam_xt}
\end{figure}
